---
title: "Role of vitamin c in tooh growth"
author: "Kevin Guillemot"
date: "April 5, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(datasets)
library(ggplot2)
library(dplyr)
library(broom)
rawData<-ToothGrowth
```

## Overview

We are studying the response is the length of odontoblasts (cells responsible for tooth growth) in 60 guinea pigs. Each animal received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods, (orange juice or ascorbic acid (a form of vitamin C and coded as VC).

## Exploratory data analysis

Length distribution seem to be unimodal and doesn't not require any scaling

```{r echo=FALSE,fig.height=2.5}
ggplot(rawData,aes(len))+
  geom_density()+
  labs(title="Length distribution")
```

Factors are in equal proportions in the dataset. Data seems to have been properly randomized.

```{r }
prop.table(table(select(rawData,supp,dose)))
```

## Influence of Supplement type

OJ seems slightly more potent than VC at first glance
```{r echo=FALSE,fig.height=3.5}
ggplot(rawData,aes(x=supp,y=len))+
  geom_boxplot()+
  labs(title="Length vs supplementC")
```

T test revleals no statistical significance. The pvalue of 6% is higher than our 5% threshold, the 95% confidence interval of supplement effect includes 0.

```{r}
tidy(t.test(formula=rawData$len~rawData$supp, var.equal = FALSE))
```

## Influence of Dose

An increase in dose seems to be related with an increase in length.

```{r echo=FALSE,fig.height=3.5}
ggplot(rawData,aes(x=dose,y=len))+
  geom_point(aes(colour = supp))+
  labs(title="Length vs dose")
```

Pairwise T tests suggest the average tooh length is different for each dose level. P avlue is smaller than the significance level of 5% in every case.

```{r}
tidy(t.test(rawData$len[rawData$dose==1],rawData$dose[rawData$dose==0.5], var.equal = FALSE))
tidy(t.test(rawData$len[rawData$dose==2],rawData$dose[rawData$dose==1], var.equal = FALSE))
tidy(t.test(rawData$len[rawData$dose==2],rawData$dose[rawData$dose==0.5], var.equal = FALSE))
```

## Conclusion

From the available dataset we can conclude tooth length is not influenced in a statistically significant way by the supplement type taken, but that it varies significantly depending on the dose administrated.

A further study could extend the range of dose tested to see if the the effect is still valid at higher dose (which we intuitively wouldn't expect).

