---
title: "Simulation on exponential distribution"
author: "Kevin Guillemot"
date: "April 5, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

## Overview

In this project you will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution rate parameter is lambda. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. 

We will then compare different sample statistics and distributions to their theoretical distributions.

## Simulation

We used R to simulate 40 exponential sample, take their mean and replicate the process 1000 times.
```{r }
# Parameters
lambda<-0.2
n<-40
nbSimulations<-1000

# Population statistics
popMean<-1/lambda
popSd<-1/lambda

# Sample mean statistics
sampleMeans<-replicate(nbSimulations,mean(rexp(n, lambda)))
meanSampleMeans<-mean(sampleMeans)
sdSampleMeans<-sd(sampleMeans)
```

## Show the sample mean and compare it to the theoretical mean of the distribution.

We observe sample mean and population mean arevery close, sample mean is an unbiased estimator of population mean.
```{r }
popMean
meanSampleMeans
```

The graph below illustrates the distribution of sample means and compares it to the the theoretical mean.

```{r echo=FALSE}
ggplot(data.frame(means=sampleMeans),aes(means))+
  geom_density() +
  geom_vline(xintercept = popMean,colour="blue")+
  geom_vline(xintercept = meanSampleMeans,colour="red")+ 
  labs(title = "Distribution of sample means")+
  annotate("text",x=c(popMean-1.5,meanSampleMeans+0.4),y=-0.01,label=c("Theoretical mean","Simulation mean"),hjust=0, colour=c("blue","red"))
```

## Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.


The variance of the sample is very close to the population variance divided by the sample size as expected by theory.
```{r }
sdSampleMeans^2
popSd^2/n
```

The graph below illustrates the distribution of sample means and compares it to the the theoretical mean.

```{r echo=FALSE}
ggplot(data.frame(means=sampleMeans),aes(means))+
  geom_density() +
  geom_vline(xintercept = popMean+popSd/sqrt(n),colour="blue")+
  geom_vline(xintercept = popMean-popSd/sqrt(n),colour="blue")+
  geom_vline(xintercept = meanSampleMeans+sdSampleMeans,colour="red")+ 
  geom_vline(xintercept = meanSampleMeans-sdSampleMeans,colour="red")+
  labs(title = "Distribution of sample means")+
  annotate("text",x=c(popMean-1.5,meanSampleMeans+1),y=-0.01,label=c("Theoretical Standard deviation","Simulation Standard deviation"),hjust=0, colour=c("blue","red"))
```

## Show that the distribution is approximately normal.

The QQ plot of the a large collection of averages of 40 exponentials is approximately normal.

```{r echo=FALSE}
y<- quantile(sampleMeans,c(0.25,0.75))
x<- qnorm(c(0.25, 0.75))
slope<-diff(y)/diff(x)
int <- y[1L] - slope * x[1L]
ggplot(data.frame(means=sampleMeans),aes(sample=means))+
  stat_qq()+
  geom_abline(slope=slope,intercept=int)+
  labs(title = "Averages of 40 exponentials")
```

A large collection of random exponentials clearly has not a normal distribution.

```{r echo=FALSE}
expSamples<-rexp(nbSimulations, lambda)
y<- quantile(expSamples,c(0.25,0.75))
x<- qnorm(c(0.25, 0.75))
slope<-diff(y)/diff(x)
int <- y[1L] - slope * x[1L]
ggplot(data.frame(s=expSamples),aes(sample=s))+
  stat_qq()+
  geom_abline(slope=slope,intercept=int)+
  labs(title = "Random exponentials")
```